\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2022-04-25 sections/results.tex
Fix: Update the captions on the figures

  Figure \ref{fig:figure2}B is the PCA plot of all the generated co-occurrence networks, colored by the databases in the TA step (left) and the network inference algorihtms in the NI step (right).
  In other words, each point corresponds to a network generated using different combination of tools, and captures how much the final network is affected by such choice.
  The importance of the TA step indicated in Figure \ref{fig:figure2}A is also reflected here, in the fact that most of the networks can be clearly separated based on the database used (Figure \ref{fig:figure2}B).
  However, it is also evident that some of the networks, especially those inferred using the direct association based network infernce methods, are much closer in the PCA plot regardless of the reference database used.
  It can also be seen that the variation (on the first two principal axes) between the networks decreases when the low abundance \ac{otu}s are removed from the network (Figure~\ref{fig:figure_s1}).
  These results suggest that the most important criterion for accurate comparative analyses of co-occurrence networks is the taxonomy reference database and the level of filtering of the taxonomy tables prior to network inference.

@@

  Denoising and clustering are commonly carried out to generate representative sequences from the demultiplexed and trimmed 16S sequencing data to obtain the \ac{otu}/\ac{esv} tables (counts of these representative sequences for each sample).
  In order to compare the representative sequence count tables generated by different tools we processed the same 16S sequencing reads (samples from a fecal microbiome transplant study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
  The first three methods are from the vsearch plugin and Deblur is from the deblur plugin of \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019}.
  The \ac{dada2} method is from the \ac{dada2} R package.

@@

  The representative sequences obtained from the \ac{dada2} method (with remove bimera for chimera checking) in the DC step were used for taxonomic assignment using the three reference databases.
  Figure~\ref{fig:figure4}A depicts a flow diagram (alluvial plot) that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three different databases.
  We observe that the top 50 representative sequences are assigned to 17 Genus in \ac{gg}, 23 in SILVA and 17 in \ac{ncbi}, however the sequences that are assigned to a particular Genus vary significantly based on the taxonomy database.
  We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (pink) also differ.
  The large percentage of unassigned sequences indicate that even among the most abundant representative sequences a large fraction is assigned an "unknown" Genus in two of the three databases (\ac{ncbi} and \ac{gg}).
  A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a Genus; the second (more likely) reason is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
  Additionally, we observe that other abundant Genus like Bacteroides and Escherichia are also assigned to different representative sequences (indicated by the different proporitions in the alluvial plot).
  Furthermore, the most abundant Genus (among the top 50 representative sequences) in the \ac{gg} taxonomy table was \textit{Escherichia} whereas it was assigned to "unknown" in NCBI and in the SILVA taxonomy table it was assigned to \textit{Escherichia-Shigella}.
  Although these are minor differences (in the case of \ac{gg} and SILVA), when comparing a large number of taxonomy composition tables these problems are hard to diagnose.

>>>>

  Figure \ref{fig:figure2}B is the PCA plot of all the generated co-occurrence networks, colored by the databases in the TA step (left) and the network inference algorithms in the NI step (right).
  In other words, each point corresponds to a network generated using different combination of tools, and captures how much the final network is affected by such choice.
  The importance of the TA step indicated in Figure \ref{fig:figure2}A is also reflected here, in the fact that most of the networks can be clearly separated based on the database used (Figure \ref{fig:figure2}B).
  However, it is also evident that some of the networks, especially those inferred using the direct association based network infernce methods, are much closer in the PCA plot regardless of the reference database used.
  It can also be seen that the variation (on the first two principal axes) between the networks decreases when the low abundance \ac{otu}s are removed from the network (Figure~\ref{fig:figure_s1}).
  These results suggest that the most important criterion for accurate comparative analyses of co-occurrence networks is the taxonomy reference database and the level of filtering of the taxonomy tables prior to network inference.

@@

  Denoising and clustering are commonly carried out to generate representative sequences from the demultiplexed and trimmed 16S sequencing data to obtain the \ac{otu}/\ac{esv} tables (counts of these representative sequences for each sample).
  In order to compare the representative sequence count tables generated by different tools we processed the same 16S sequencing reads (samples from a fecal microbiome transplant study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
  The first three methods are from the vsearch plugin and Deblur is from the deblur plugin of \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019}.
  The \ac{dada2} method is from the \ac{dada2} R package.
  The closed and open reference methods in this analysis use the \ac{gg} database for reference sequence alignment.

@@

  The representative sequences obtained from the \ac{dada2} method (with remove bimera for chimera checking) in the DC step were used for taxonomic assignment using the three reference databases.
  Figure~\ref{fig:figure4}A depicts a flow diagram (alluvial plot) that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three different databases.
  We observe that the top 50 representative sequences are assigned to 17 Genus in \ac{gg}, 23 in SILVA and 17 in \ac{ncbi}, however the sequences that are assigned to a particular Genus vary significantly based on the taxonomy database.
  We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (pink) also differ.
  The large percentage of unassigned sequences indicate that even among the most abundant representative sequences a large fraction is assigned an "unknown" Genus in two of the three databases (\ac{ncbi} and \ac{gg}).
  A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a Genus; the second (more likely) reason is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
  Additionally, we observe that other abundant Genus like Bacteroides and Escherichia are also assigned to different representative sequences (indicated by the different proporitions in the alluvial plot).
  Furthermore, the most abundant Genus (among the top 50 representative sequences) in the \ac{ncbi} taxonomy table was \textit{Escherichia} whereas it was assigned to "unknown" in \ac{gg} and in the SILVA taxonomy table it was assigned to \textit{Escherichia-Shigella}.
  Although these are minor differences (in the case of \ac{ncbi} and SILVA), when comparing a large number of taxonomy composition tables these problems are hard to diagnose.

### 2022-04-25 sections/tables_figures.tex
Add: FMT figures to manuscript
Change: Update discussion section
Fix: Update the captions on the figures

  \begin{table}[h]
    \centering
    \small
    \begin{tabular}{|c|c|c|}
      \hline
      \textbf{Process} & \textbf{Tool} & \textbf{Parameters} \\
      \hline
      Denoising and Clustering & Dada2/Deblur & default \\
      Taxonomy Assignment & \ac{ncbi} with Blast & RefSeq database \\
      OTU Processing & Based on statistical power & Dynamic cutoff \\
      Network Inference & Consensus method & - \\
      \hline
    \end{tabular}
    \caption{Default tools and parameters for the pipeline}
    \label{tab:default_options}
  \end{table}

@@

  \begin{table}[h]
  \begin{tabular}{|c|c|c|}
  \hline
  Condition                    & Parameter                 & Value          \\ \hline
  \multirow{3}{*}{Filter(off)} & Prevalence threshold      & 2 / n\_samples \\ \cline{2-3}
                               & Abundance threshold       & 0.001          \\ \cline{2-3}
                               & Observation sum threshold & 10             \\ \hline
  \multirow{3}{*}{Filter(on)}  & Prevalence threshold      & 0.05           \\ \cline{2-3}
                               & Abundance threshold       & 0.01           \\ \cline{2-3}
                               & Observation sum threshold & 100            \\ \hline
  \end{tabular}
  \caption{Parameters used in the OP module}
  \label{tab:ni_tools}
  \end{table}

@@

% Figures

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.75\linewidth]{figure1.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
      The steps can be grouped into four major groups: \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU or ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
      Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
      The text along the arrows describes the data that is being passed from one step to another. For details on each process and data types, see Methods.
    }
    \label{fig:figure1}
  \end{figure}

@@

  \begin{figure}[H]
    \centering
    \includegraphics[width=1.0\linewidth]{figure2.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
      \caption{
      \textbf{The choice of database contributes to the most variance in the networks}.
      \textbf{(A)} The total relative variance in the networks contributed by the DC, TA and OP steps of the pipeline (right) and the linear model used to calculate the relative variance (left), see the Methods section for details.
      \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
      The color of the points corresponds to the taxonomy database, the shape corresponds to the denoising/clustering method and the size corresponds to whether low abundance OTUs were removed or not.
      \textbf{(B inset)} The network generated using DC=dada2, TA=GG, OP=no and NI=SPARCC and represents the particular point shown (big red square).
      The points on the PCA plot can be separated based on the TA step and the differences due to the DC and OP steps are comparatively not as significant.
    }
    \label{fig:figure2}
  \end{figure}
  \FloatBarrier
  \newpage

  % TODO: Describe the datasets being shown here clearly. In the caption + figure legends (if necessary) and in the text
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figure3.pdf}
    \caption{
      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are in low abundance.}
      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods,
      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences of low abundance,
      \textbf{(C, D)} The distributions of the average weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets.
      \textbf{(D)} The distributions of the average unweighted UniFrac distance show that dada2 and Deblur were the best performing methods in most of the datasets.
    }
    \label{fig:figure3}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth,height=1.2\textwidth]{figure4.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments.}
      \textbf{(A)} The assignment of the top 50 representative sequences to their respective taxonomies using the three different reference databases shows how the same sequences are assigned to different Genus.
      \textbf{(B)} The percentage of \ac{otu}s assigned to the same taxonomic label when using different reference databases.
      The percentage of mismatches decrease at higher taxonomic levels but even at the Phylum level there exists around 10\% of mismatches.
      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset.
    }
    \label{fig:figure4}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{figure5.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
      \textbf{(A)} The six different networks generated by the different network inference methods are very dissimilar.
      The green links are positive associations and the orange links are negative associations.
      A threshold of 0.3 was set for the methods that infer pairwise correlations (\ac{sparcc}, Spearman, Pearson) and no threshold was set for the other methods.
      \textbf{(B)} The node overlap Upset plot~\cite{Lex} indicates that all the networks have a large number of common nodes involved in connections.
      Whereas, \textbf{(C)} The edge overlap Upset plot shows that a very small fraction of these connections are actually shared.
    }
    \label{fig:figure5}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[h]
    \centering
    \includegraphics[width=1.0\linewidth]{figure6.pdf}
    \caption{
      The precision and sensitivity of the inferred networks on the synthetic data.
    }
    \label{fig:figure6}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[h]
    \centering
    \includegraphics[width=0.85\linewidth]{figure7.pdf}
    \caption{
      \textbf{Network inference and taxonomic assignment have the highest influence on the inferred network structures.}
      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks generated when one of the steps use a different tool.
      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network but not present in the current network are gray.
      \textbf{(B)} The L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
    }
    \label{fig:figure7}
  \end{figure}

>>>>

  \begin{table}[h]
    \centering
    \small
    \begin{tabular}{|c|c|c|}
      \hline
      \textbf{Process} & \textbf{Tool} \\
      \hline
      Denoising and Clustering & Dada2 and remove bimera \\
      Taxonomy Assignment & \ac{gg} with NaiveBayes classifier \\
      OTU Processing & Thresholds on prevalence and abundance \\
      Network Inference & scaled-sum consensus method \\
      \hline
    \end{tabular}
    \caption{Default tools and parameters for the pipeline}
    \label{tab:default_options}
  \end{table}

@@

  \begin{table}[h]
    \centering
    \small
    \begin{tabular}{|c|c|c|}
    \hline
    Condition                    & Parameter                 & Value          \\ \hline
    \multirow{3}{*}{Filter(off)} & Prevalence threshold      & 2 / n\_samples \\ \cline{2-3}
                                 & Abundance threshold       & 0.001          \\ \cline{2-3}
                                 & Observation sum threshold & 10             \\ \hline
    \multirow{3}{*}{Filter(on)}  & Prevalence threshold      & 0.05           \\ \cline{2-3}
                                 & Abundance threshold       & 0.01           \\ \cline{2-3}
                                 & Observation sum threshold & 100            \\ \hline
    \end{tabular}
    \caption{Parameters used in the OP module}
    \label{tab:ni_tools}
  \end{table}

@@

% Figures
  \FloatBarrier
  \newpage

  \begin{figure}
    \centering
    \includegraphics[width=0.75\linewidth]{figure1.pdf}
  \end{figure}
  \begin{figure}[h]
    \centering
    \caption{
      \textbf{The workflow of the \ac{micone} pipeline}.
      The steps can be grouped into five major groups: \textbf{(SP)} \textbf{S}equence \textbf{P}rocessig, \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU or ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
      Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
      The text along the arrows describes the data that is being passed from one step to another.
      The final output of the pipeline are consensus networks generated from the co-occurrence networks.
      For details on each process and data types, see Methods.
    }
    \label{fig:figure1}
  \end{figure}

@@

  \begin{figure}
    \centering
    \includegraphics[width=1.0\linewidth]{figure2.pdf}
  \end{figure}
  \begin{figure}
    \centering
      \caption{
      \textbf{The choice of reference database contributes to the most variance in the networks}.
      \textbf{(A)} The total relative variance in the networks contributed by the DC, CC (chimera checking), TA and OP steps of the pipeline (right) and the linear model used to calculate the relative variance (left), see the Methods section for details.
      The analyses was performed on the healthy stool samples in a fecal microbiome transplant dataset~\ref{Kang2017}.
      The taxonomy database contributes most to the variance between the networks (65.4\%) followed by filtering of the counts matrix (26.8\%) in the OP step.
      \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
      Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
      The color of the points corresponds to the reference taxonomy database used to reconstruct those networks (left), and the network inference algorithm used to infer the networks (right).
      % TODO: The points aren't really separated
      The points on the PCA plot can be separated based on the TA step and the differences due to the DC and CC steps are comparatively not as significant.
    }
    \label{fig:figure2}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figure3.pdf}
    \caption{
      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are lower in abundance.}
      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB)).
      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the calculated sequence profile in the mock datasets.
      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
      The above average performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
    }
    \label{fig:figure3}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth,height=1.2\textwidth]{figure4.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments beyond the Order level.}
      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases show how the same sequences are assigned to different Genus under the different databases.
      A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
      The number of mismatches are fewer at higher taxonomic levels but even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
      However, we recommend the \ac{gg} database due to its popularity and if the goal is to compare against existing networks in literature.
    }
    \label{fig:figure4}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{figure5.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
      \textbf{(A)} The nine different networks generated by the different network inference methods (excluding \ac{mldm}) are very dissimilar.
      The green links are positive associations and the orange links are negative associations.
      A threshold of 0.3 was set for the correlation-based methods (\ac{sparcc}, Propr, Spearman and Pearson) and a threshold of 0.01 was set for the other direct association methods.
      \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large number of common nodes involved in connections (33 out of 68).
      Whereas, \textbf{(C)} the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
    }
    \label{fig:figure5}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[h]
    \centering
    \includegraphics[width=1.0\linewidth]{figure6.pdf}
    \caption{
      \textbf{The precision and sensitivity of the inferred networks on the NorTA synthetic interaction data.}
      The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
      The different consensus based methods used are: pvalue merging, scaled-sum (SS) and simple voting (SV) method.
      Pearson and Spearman methods are not used in the calculation of the consensus.
      Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.944), but the overall best precision was consistently obtained by the scaled-sum method (0.956, 0.985 and 1.000).
      The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), also outperforms \ac{spieceasi} on average precision (0.969).
      Therefore, we recommend the scaled-sum consensus method as the default tool for network inference.
    }
    \label{fig:figure6}
  \end{figure}
  \FloatBarrier
  \newpage

  % TODO: Verify the validity of Figure 7A and 7B again
  % Are they both using consensus?
  \begin{figure}[h]
    \centering
    \includegraphics[width=0.85\linewidth]{figure7.pdf}
    \caption{
      \textbf{The consensus algorithm drastically reduces variability in the networks and increases robustness of inferred networks}
      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, CC=remove bimera, TA=\ac{gg}, OP=Filter(on), NI=scaled-sum consensus) is compared with networks generated when one of the steps use a different tool.
      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network but not present in the current network are gray.
      Changing the NI, TA and OP steps leads to the creation of the most number of new nodes and edges.
      \textbf{(B)} The boxplot showing the L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
      The high value of L1 distances for the network inference step as compared to the other steps demonstrates how the consensus method is able to reduce the variability in the networks in the other steps of the pipeline.
      \textbf{(C)} The networks generated for the control (left) and autistic (right) samples in the fecal microbiome transplant dataset.
      These networks were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
    }
    \label{fig:figure7}
  \end{figure}
